---
title: "Explore climate-production relationships"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Links between climate and maple syrup production 

Maple syrup production is intrinsically linked to climate due to known biophysical processes that result in positive stem pressures after freeze-thaw cycles [@REF]. Several previous studies showed general links between preceding meteorological conditions and syrup production [@REF], as well as the underlying production volume and sugar content [@REF]. 

We will explore the climatic relationships between production data from the *Producteurs et Productrices Acéricole du Québec* (PPAQ) at the municipal scale for the 2005 to 2021 period and climate data simulated for each producing municipality. We simulated the following climate variables using BioSIM [@REF] for every day from 2004 to 2021:

* Minimum, mean and maximum temperature
* Total rainfall, snowfall, and precipitation
* Mean snow cover
* Mean relative humidity
* Mean wind direction and speed (at 2 and 10 m)
* Total down-welling shortwave radiation
* Mean atmospheric pressure 

### Biologically meaningful variables



```{r sourceData, echo = FALSE}
# load matched climate and production data
if (!exists ("production") source ("matchClimAndProdData.R"))

```
