---
title: "Explore climate-production relationships"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load dependencies
#----------------------------------------------------------------------------------------
library ("lme4") # for mixed effects models 
library ("brms") # for bayesian models
library ("readr") # to read csv files

# load matched climate and production data
#----------------------------------------------------------------------------------------
if (!exists ("production")) production <- read_csv ("../data/productionData.csv") 

# to-do list:
#----------------------------------------------------------------------------------------
# TR - Add bibliography and citations 
# TR - Insert graphic of periodsOfInfluence and write section about derived variables
# TR - Integrate more realistic models to explain 
```

## Links between climate and maple syrup production 

Maple syrup production is intrinsically linked to climate due to known biophysical processes that result in positive stem pressures after freeze-thaw cycles [@REF]. Several previous studies showed general links between preceding meteorological conditions and syrup production [@REF], as well as the underlying production volume and sugar content [@REF]. 

We will explore the climatic relationships between production data from the *Producteurs et Productrices Acéricole du Québec* (PPAQ) at the municipal scale for the 2005 to 2021 period and climate data simulated for each producing municipality. We simulated the following climate variables using BioSIM [@REF] for every day from 2004 to 2021:

* Minimum, mean and maximum temperature
* Total rainfall, snowfall, and precipitation
* Mean snow cover
* Mean relative humidity
* Mean wind direction and speed (at 2 and 10 m)
* Total down-welling shortwave radiation
* Mean atmospheric pressure 

### Biologically meaningful variables

Let's start with looking at the climate space we are covering.   

```{r climateSpacePlot, echo = FALSE}
# plot climate space of the production in Quebec
#----------------------------------------------------------------------------------------
plot (x = production$tMean,
      y = production$pTota,
      xlab = expression (paste ("Mean annual temperature (",degree,"C)")),
      ylab = "Total annual precipitation (mm)", pch = 19, col = "#cc724011",
      axes = FALSE, xlim = c (0, 10), ylim = c (0, 1800))
axis (side = 1, at = seq (0, 10, by = 2))
axis (side = 2, at = seq (0, 1800, by = 300), las = 1)
```

Next, we can look at any relationship between mean annual temperature and total annual precipitation versus production per tap ($L \, tap^{-1}$).


```{r tMeanVmeanP, echo = FALSE}
par (mfrow  = c (1, 2))
# plot climate space of the production in Quebec
#----------------------------------------------------------------------------------------
plot (x = production$tMean,
      y = production$meanP,
      xlab = expression (paste ("Mean annual temperature (",degree,"C)")),
      ylab = expression (paste ("Approximated maple yield (L ", tap^-1,")")), 
      pch = 19, col = "#cc724011",
      axes = FALSE, xlim = c (0, 10), ylim = c (0, 5))
axis (side = 1, at = seq (0, 10, by = 2))
axis (side = 2, at = 1:5, las = 1)
plot (x = production$pTota,
      y = production$meanP,
      xlab = "Total annual precipitation (mm)",
      ylab = expression (paste ("Approximated maple yield (L ", tap^-1,")")), 
      pch = 19, col = "#cc724011",
      axes = FALSE, xlim = c (0, 1800), ylim = c (0, 5))
axis (side = 1, at = seq (0, 1800, by = 300))
axis (side = 2, at = 1:5, las = 1)
```